### 2024-05-14 thesis/chapters/03-design.tex
Move Resume to main thesis

\chapter{Design} \label{chapter:design}
During the design stage, we set out research questions and detail the steps of the machine learning pipeline for diagnosing fault types in the MaFaulDa dataset. The explanatory analysis of MaFaulDa hints at the underlying dependencies within feature sets that can degrade the machine learning model of choice. We apply the methodology for vibration measurements from standards to the data logger requirements and data acquisition from industrial equipment.

\section{Research questions}
The thesis aims to provide answers to five research questions. The main focus is on making data flow more efficient in an industrial sensor network that monitors rotating machines. The \textbf{research questions} are:
\begin{enumerate}[label=RQ\arabic*., font=\bfseries]
    \itemsep0pt
	\item Which time-domain and frequency-domain features are extractable from vibrations to provide the most accurate record of machinery faults?
	\item What is the reduction in transmission goodput with signal features?
	\item What accuracies of fault classification are achievable with tiny feature subsets?
	\item How do the vibration signals collected on the industrial machinery behave?
	\item How can machinery faults be identified over time from collected events?
\end{enumerate}

\noindent In accomplishing the objectives of our research, we propose several \textbf{goals}:
\begin{todolist}
    \itemsep0pt
    \item Statistically and visually describe vibration signals from the Machinery fault database (MauFaulDa).
    \item Establish a list of conditions that will be investigated later in the experimental setting.
    \item Prepare the dataset to be used in conjunction with machine learning models namely by identifying labels and balancing classes.
    \item Find the best subsets of features in the time and frequency domain with previously analyzed feature extraction and selection methods.
    \item Evaluate the performance of models described in the diagnostics section with a significant focus on the k-nearest neighbor algorithm.
    \item Acquire measurements of vibrations from machines in the real environment to form a novel dataset of machinery behavior.
    \item Develop hardware and implement its firmware to obtain such measurements of the quality demanded by vibrodiagnostics standards.
\end{todolist}

\section{Machine learning pipeline}
The similar data processing framework we adopt can be found elsewhere in the literature. It generally consists of signal acquisition, feature extraction, dimensionality reduction, and pattern recognition or fault detection~\cite{wang_bearing_2015}. However, the realization of the steps must be decomposed closer to suit the dataset structure, desired processing goals, and investigated parameters. The entire design of the data flow can be seen in Fig.~\ref{fig:design:ml-pipeline}.

@@

\textbf{In preprocessing step}, the signals stored in columns of CSV files within the MaFaulDa ZIP archive are first associated with metadata according to the file's path in the directory structure. The hierarchically topmost folder describes the simulated defect. There is directory for \emph{normal}, \emph{imbalance}, \emph{horizontal-misalignment}, \emph{vertical-misalignment}, \emph{overhang} (outer bearing), \emph{underhang} (inner bearing). Each bearing has an additional folder layer for dividing bearing faults: \emph{ball\_fault}, \emph{cage\_fault}, \emph{outer\_race}. Each fault category except the normal class contains various fault severities with units of grams or millimeters. The severities in turn consist of individual files with different rpm speeds.

We neglect the \textbf{acceleration sensor orientation}, therefore, the labels for shaft misalignment in vertical and horizontal directions are merged into the same group. In the end, that leaves \textbf{6 types of labels}: baseline, two shaft faults are imbalance and misalignment, and three bearing faults are cage fault, ball fault, and outer race fault.

Depending on the \textbf{chosen bearing position}, only time series associated with that bearing are retained because the standards demand each bearing be assessed separately. Fault classification concerns bearing in direct contact and shaft mechanically passing through it.

@@

\textbf{File description} is the triplet of \textbf{fault}, \textbf{severity}, and \textbf{rpm}.

After labeling we insert a step to split the time series into parts when the duration is long enough to introduce more observations artificially. This assumption for splitting is not met for MaFaulDa because, for the desired frequency resolution for features of close to 1~Hz, we need a window size of 32768 samples. Variation reduction in spectral estimation brings the necessity to average at least 12 overlapping windows. The full length of 5 seconds provides 15 whole overlapping windows.

@@

The preprocessed signal from each axis of accelerometer is packed up to \textbf{two base feature sets} in time domain (TD) and frequency domain (FD). The feature extraction formulas are is identical with formulas presented in the section \ref{section:feature-extraction} of analysis chapter. The frequency transform is Welch's method averaging FFT vectors from 32768 samples after Hann windowing and 50\% overlap. The feature sets are:
\begin{itemize}
\itemsep0pt
\item \textbf{TD has 10 time-domain features} - peak-to-peak amplitude (\emph{pp}), zero-crossing rate (\emph{zerocross}), root mean square (\emph{rms}), skewness, kurtosis, shape factor, crest factor, impulse factor, clearance factor, average amplitude change (aac), 
\item \textbf{FD has 11 frequency-domain features} - spectral centroid, standard deviation (\emph{std}), skewness, kurtosis, roll-on frequency, roll-off frequency, spectral flux, noisiness, spectral negentropy, energy, entropy.
\end{itemize}

The four conditions are applied in the instance of the pipeline to filter observations with calculated features to create \textbf{24 scenarios}:
\begin{itemize}
\itemsep0pt
\item \textbf{Source bearings} - samples and fault labels are left in either just for the inner bearing (A), for only the outer bearing (B), or for both bearings (A+B).
\item \textbf{Feature domain} - the set of extracted features as an input to downstream models is changed either to TD or FD set,
\item \textbf{Accelerometer axis} - Either one principal direction of motion or all three dimensions is aggregated under same feature name (1 or 3),
\item \textbf{High severity faults} - to more precisely simulate fault frequencies found in real environment the low severity defects can be considered as normal operation (Yes or No).
\end{itemize}

>>>>

\chapter{Design} \label{chapter:design}
During the design stage, we set out research questions and explain the machine learning pipeline for machinery defects diagnostics in the MaFaulDa dataset. The explanatory analysis of MaFaulDa hints at the underlying dependencies within feature sets that can degrade the machine learning model of choice. We apply the methodology for vibration measurements found in technical standards to requirements for the data logger and data acquisition procedures for the industrial equipment.

\section{Research questions}
The thesis aims to provide answers to five research questions. The main focus is on making data flow more efficient in an industrial sensor network that could monitor rotating machines. The \textbf{research questions} are:
\begin{enumerate}[label=RQ\arabic*., font=\bfseries]
    \itemsep0pt
	\item Which time-domain and frequency-domain features are extractable from vibrations to provide the most accurate record of machinery faults?
	\item What is the reduction in transmission goodput with signal features?
	\item What accuracies of fault classification are achievable with tiny feature subsets?
	\item How do the vibration signals collected on the industrial machinery behave?
	\item How can machinery faults be identified incrementally from collected events?
\end{enumerate}

\noindent In accomplishing the objectives of our research, we propose several \textbf{goals}:
\begin{todolist}
    \itemsep0pt
    \item Statistically and visually describe vibration signals from the Machinery fault database (MauFaulDa).
    \item Establish a list of conditions that will be investigated in the experiments.
    \item Prepare the dataset to be used in conjunction with machine learning models namely by identifying labels and balancing classes.
    \item Find the best subsets of features in the time and frequency domain using previously analyzed feature extraction and selection methods.
    \item Evaluate the performance of models described in the diagnostics section with a significant focus on the k-nearest neighbours algorithm.
    \item Acquire measurements of vibrations from machines in the real environment to form a novel dataset of machinery behavior.
    \item Develop hardware for a sensor unit and implement its firmware to obtain vibration measurements with the quality demanded by vibrodiagnostics standards.
\end{todolist}

\section{Machine learning pipeline}
We adopt a similar data processing framework found also elsewhere in the literature. It generally consists of signal acquisition, feature extraction, dimensionality reduction, and pattern recognition or fault detection~\cite{wang_bearing_2015}. However, the realization of the steps must be decomposed closer to suit the dataset structure, desired processing goals, and investigated parameters. The entire design of the data flow is in Figure~\ref{fig:design:ml-pipeline}.

@@

The signals are in columns of CSV files within the MaFaulDa ZIP archive. \textbf{In preprocessing step}, they are first associated with metadata according to their file's path in the directory structure. The hierarchically topmost folder describes the simulated defect. There is directory for \emph{normal}, \emph{imbalance}, \emph{horizontal-misalignment}, \emph{vertical-misalignment}, \emph{overhang} (outer bearing), \emph{underhang} (inner bearing). Each bearing has an additional folder layer for dividing up bearing faults: \emph{ball\_fault}, \emph{cage\_fault}, \emph{outer\_race}. Fault categories, except the normal class, contain various fault severities marked with units of grams or millimeters. Severities, in turn, consist of individual files with different rpm speeds.

The labels for shaft misalignment in vertical and horizontal directions are merged into the same group because we neglect the \textbf{acceleration sensor orientation}. In the end, that leaves \textbf{six types of labels}: baseline, two shaft faults are imbalance and misalignment, and three bearing faults are cage fault, ball fault, and outer race fault.

Depending on the \textbf{chosen bearing position}, only time series associated with that bearing are retained because the standards demand each bearing be assessed separately. Fault classification concerns with bearing in direct contact and shaft mechanically passing through it.

@@

\textbf{File description} is the triplet of \textbf{fault}, \textbf{severity}, and \textbf{rpm}. After labeling, we insert a step to split the time series into parts when the duration is long enough to introduce more observations artificially. This assumption for splitting is not met for MaFaulDa because, for the desired frequency resolution for features of close to 1~Hz, we need a window size of 32768 samples. Variation reduction in spectral estimation necessitates averaging at least twelve overlapping windows. The full length of five seconds provides fifteen complete windows.

@@

The preprocessed signal from each axis of accelerometer is packed up to \textbf{two base feature sets} in time domain (TD) and frequency domain (FD). The feature extraction formulas are identical with formulas presented in the section \ref{section:feature-extraction} of analysis chapter. The frequency transform is Welch's method averaging of FFT vectors from 32768 samples after Hann windowing and 50\% overlap. The feature sets are:
\begin{itemize}
\itemsep0pt
\item \textbf{TD has 10 time-domain features} - peak-to-peak amplitude (\emph{pp}), zero-crossing rate (\emph{zerocross}), root mean square (\emph{rms}), skewness, kurtosis, shape factor, crest factor, impulse factor, clearance factor, average amplitude change (aac), 
\item \textbf{FD has 11 frequency-domain features} - spectral centroid, standard deviation (\emph{std}), skewness, kurtosis, roll-on frequency, roll-off frequency, spectral flux, noisiness, spectral negentropy, energy, entropy.
\end{itemize}

The four conditions are applied in each instance of the pipeline to filter observations with calculated features to create \textbf{24 scenarios}:
\begin{itemize}
\itemsep0pt
\item \textbf{Source bearings} - samples and fault labels are left in either just for the inner bearing (A), for only the outer bearing (B), or for both bearings (A+B).
\item \textbf{Feature domain} - the set of extracted features as an input to downstream models is changed either to TD or FD set,
\item \textbf{Accelerometer axis} - Either one principal direction of motion or all three dimensions are aggregated under same feature name (1 or 3),
\item \textbf{High severity faults} - to more precisely simulate fault frequencies found in real environment the low severity defects can be considered as normal operation (Yes or No).
\end{itemize}

### 2024-05-14 thesis/chapters/04-implementation.tex
Move Resume to main thesis

\chapter{Implementation} \label{chapter:implementation}
The tools implemented include exploration of datasets using statistical overviews and visualizations, experiments of feature selection which involve machine learning pipeline, and firmware for data logger's hardware.

@@

\section{Firmware}
The drivers for low-level interfaces of ESP32 microcontroller and FAT32 filesystem are already available in Espressif ESP-IDF SDK. The accelerometer driver is made available by the vendor\footnote{IIS3DWB driver: \url{https://github.com/STMicroelectronics/iis3dwb-pid}}. The entry routine of the firmware mounts the SD card, sets up GPIO pins for LED and button, and executes three \emph{FreeRTOS} tasks. Procedures in firmware are described in Appendix~\ref{appendix:technical-docs}. The hardware (Fig.~\ref{fig:hw-data-logger}) was built by the thesis consultant based on the complete specification supplied by the author.

@@

Another impediment presented the slow string formatting with \emph{printf} family of functions during the accelerometer sampling. The readings cannot be written to TSV directly because the real-time (RT) deadline is not met due to the high sampling rate. It is true even if the queue is utilized and the processing takes place in another task. The binary files from the SD card are converted in bulk to TSV with custom Python command line script \verb|bin2tsv.py|.

During debugging, the pin was set to change the output level immediately before the one buffer started file write and after its completion. The scopes in Figure~\ref{fig:implementation:file-write} show the time the write took for binary compared to the TSV format. In writing to TSV format, the relation of numbers of already processed buffers $x$ to those left to process $y$ is: $y = 3.7x$. It is a monotonically increasing function, so the system would eventually overrun any sized assigned queue.

>>>>

\chapter{Implementation} \label{chapter:implementation}
The tools we implement include the exploration of datasets using statistical overviews and visualizations, feature selection experiments that involve a machine learning pipeline, and firmware for the data logger's hardware.

@@

\section{Firmware}
The drivers for low-level interfaces of ESP32 microcontroller and FAT32 filesystem are already available in Espressif ESP-IDF SDK. The accelerometer driver is made available by the vendor\footnote{IIS3DWB driver: \url{https://github.com/STMicroelectronics/iis3dwb-pid}}. The entry routine of the firmware mounts the SD card, sets up GPIO pins for LED and button, and executes three \emph{FreeRTOS} tasks. Procedures in firmware are described in Appendix~\ref{appendix:technical-docs}. The hardware (Fig.~\ref{fig:hw-data-logger}) was built by the thesis consultant based on the detailed specification supplied by the author.

@@

Slow string formatting with \emph{printf} family of functions during the accelerometer sampling presented another impediment. The readings cannot be written to TSV directly because the real-time (RT) deadline is not met due to the high sampling rate. This is true even if the queue is utilized and the processing takes place in another task. The binary files from the SD card are converted in bulk to TSV with custom Python script \verb|bin2tsv.py|.

During debugging, the pin was set to change the output level immediately before the one buffer started file write and after its completion. Waveforms in Figure~\ref{fig:implementation:file-write} show the time it took to write in binary format compared to the TSV format. The relation of already processed buffers $x$ to those left to process $y$ is $y = 3.7x$ for TSV format. It is a monotonically increasing function, so the system would eventually overrun queue of any size.

### 2024-05-14 thesis/chapters/05-evaluation.tex
Move Resume to main thesis

A significant change in the data stream occurs after 294 out of 1438 observations (or after 737 for high severity faults) when all 49 (or 737) normal conditions are consumed in the training process. Counters of other faults show that predictions are skewed towards more represented imbalance and misalignment. The uneven evolution of categories in a stream impacts the development of accuracy in the remaining experiments. The test accuracies of comparable batch models from the three best features are 85.47\% (TD), 87.52\% (FD), 91.71\% (TD severity), and 91.94\% (FD severity).

@@

\begin{figure}[]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-TD.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-FD.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-TD-severity.png}
        \caption{Time-domain features (severity)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-FD-severity.png}
        \caption{Frequency-domain features (severity)}
    \end{subfigure} 
    \caption{Omission of labels during incremental learning with tumbing window of length 10 and gaps of size 0, 10, and 50 samples}
    \label{fig:evaluation:label-skips-online}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/motors.png}
        \caption{Motors}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/motors-dB.png}
        \caption{Pumps}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/pumps.png}
        \caption{Motors}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/pumps-dB.png}
        \caption{Pumps dB}
    \end{subfigure}
    \caption{Characteric bearing frequencies of pumps}
    \label{fig:evaluation:bearing-freq}
\end{figure}

Domain experts recommended the procedure for fault identification by calculation of bearing characteristic frequencies (Tab.~\ref{tab:evaluation:bearing-freq-pump}), and they approved the following results. Figure~\ref{fig:evaluation:bearing-freq} identifies harmonics of rotational speed and BPFO frequency in every machine and BPFI for M2-2. It can be assumed that those frequencies will be the reason for their damage in the future. The absolute acceleration is minuscule in current frequency bands and year-long sensor log of rms velocity (Fig~\ref{fig:evaluation:ksb-guard-rms-vibartions}).

>>>>

A significant change in the data stream occurs after 294 out of 1438 observations (or after 737 for high severity faults) when all 49 (or 737) normal conditions are consumed in the training process. Counters of other faults show that predictions are skewed towards more represented classes of imbalance and misalignment. The uneven evolution of categories in a stream impacts the development of accuracy in the remaining experiments. The test accuracies of comparable batch models from the three best features are 85.47\% (TD), 87.52\% (FD), 91.71\% (TD severity), and 91.94\% (FD severity).

@@

\begin{figure}[]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-TD.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-FD.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-TD-severity.png}
        \caption{Time-domain features (severity)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-FD-severity.png}
        \caption{Frequency-domain features (severity)}
    \end{subfigure} 
    \caption{Omission of labels during incremental learning with tumbling window of length 10 and gaps of size 0, 10, and 50 samples}
    \label{fig:evaluation:label-skips-online}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/motors.png}
        \caption{Motors}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/motors-dB.png}
        \caption{Motors (dB)}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/pumps.png}
        \caption{Pumps}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/pumps-dB.png}
        \caption{Pumps (dB)}
    \end{subfigure}
    \caption{Characteric bearing frequencies of pumps}
    \label{fig:evaluation:bearing-freq}
\end{figure}

Domain experts recommended the procedure for fault identification by calculation of bearing characteristic frequencies (Tab.~\ref{tab:evaluation:bearing-freq-pump}), and they approved the following results. Figure~\ref{fig:evaluation:bearing-freq} identifies harmonics of rotational speed and BPFO frequency in every machine and BPFI for M2-2. It can be assumed that those frequencies will be the reason for damage in the future. The absolute acceleration is minuscule in current frequency bands and year-long sensor log of rms velocity (Fig~\ref{fig:evaluation:ksb-guard-rms-vibartions}).

### 2024-05-14 thesis/chapters/06-conclusion.tex
Move Resume to main thesis

The groups of three features are chosen in each domain to achieve even more pronounced data savings. We strive to keep the predictors that together have the most resemblance to the target variable. Feature selection metrics of the correlation coefficient, F statistic, mutual information, and their rank product are computed in supervised learning scenarios. The features are squished from multiple dimensions using the Euclidian norm. Lossy compression ratios attained are 2381:1 for all features and 25000:1 for six features from the MaFaulDa dataset. We managed to discard more than 99.995\% of the irrelevant data (\textbf{RQ2}).

The designed preprocessing pipeline creates the attributes from time series. The k-nearest neighbour classifier ascertains their relative fault detection power by five-fold cross-validation. The experiments for k-NN model evaluation include complete feature sets with one for each domain, the exhaustively enumerated feature subset combinations, and features subset chosen by filtering feature selection methods.

We found that the time-domain features reach higher accuracies than the frequency-domain set because of less intercorrelations. Increase in the number of neighbours for k-NN used on complete sets, the smaller the model accuracy. Above the k-value of thirteen, the trend for complete feature sets flattens out gradually. Predictors aggregated from tri-axis recording performed better in classification than those from one axis.

The k-NN accuracy with different feature subset combinations pulled out of the complete sets has a standard deviation of 7\%, and it is 10\% within IQR. Within such range, the feature selection methods can pick, in most cases, a group with accuracy above the upper quartile of accuracy distribution of every attribute combination with the same cardinality. Training k-NN on three principal components produced worse accuracy than filtering feature selection.

The ensemble feature selection methods as the rank product are necessary to balance the adverse effect of individual selection metrics. In the slight majority of 44\% of 216 dataset alterations, the rank product found the best-performing attributes closely followed by mutual information. The rank product also reaches the 97.50\% percentile in the accuracy distribution in the same-sized sets. Mutual information has a better distribution percentile of 2.84\% from rank product \textbf{(RQ3)}.

The number of features has a direct proportionality effect accuracy but at most 6\% increase is observed in moving from two to three features. The absolute best predictor triplets for MaFaulDa are in TD (accuracy of 85.47\%): zero-crossing rate, peak-to-peak, skewness, and in FD (accuracy of 87.52\%): spectral centroid, roll-off frequency, and entropy \textbf{(RQ3)}.

@@

Collected dataset analysis confirmed the stationarity of the signal spectrum of the signal and length of five-second time series as satisfactory for feature extraction. The dataset serves as a practical demonstartion of labeling potentialy future faults from bearing defect frequency amplitudes. It also proved the immense difficulty of obtaining observations of faulty machinery parts in the field without purposefully damaging bearings and offsetting the shaft \textbf{(RQ4)}.

@@

\textbf{Future work} can extend various aspects and additional questions arising from the experiments in the thesis. The natural extension is to confirm that the results hold in other publicly available datasets of machinery faults, e.g. in CWRU. The other alteration is to use more different starting base feature sets. Harmonic frequencies or wavelet coefficients can generate predictors that could be further reduced to the smallest possible subsets.

Ensemble strategies inspired by classical election systems may be utilized in the task of feature selection. Different feasible types of classifiers for online learning could be compared to k-nearest neighbours, e.g. Hoeffding tree. A data logger could be combined with the classifier model and a mechanism to annotate observation either automatically depending on the overall vibration level or manually during maintenance or after an unexpected failure.

The strategic aim would be to connect sensors on a large scale on multiple test rigs to train individual models on each bearing and combine them in a federative learning approach. Of course, vibration monitoring can be employed in other diagnostic tasks with no direct expert knowledge, such as train wheel or rail degradation due to wear and tear.

>>>>

The groups of three features are chosen in each domain to achieve even more pronounced data savings. We strive to keep the predictors that together have the most resemblance to the target variable. Feature selection metrics of the correlation coefficient, F statistic, mutual information, and their rank product are computed in supervised learning scenarios. The features are squished from multiple dimensions using the Euclidean norm. Lossy compression ratios attained are 2381:1 for all features and 25000:1 for six features from the MaFaulDa dataset. We managed to discard more than 99.995\% of the irrelevant data (\textbf{RQ2}).

The designed preprocessing pipeline creates the attributes from time series. The k-nearest neighbour classifier ascertains their relative fault detection power by five-fold cross-validation. The experiments for k-NN model evaluation include complete feature sets with one for each domain, the exhaustively enumerated feature subset combinations, and feature subsets chosen by filtering feature selection methods.

We found that the chosen time-domain features have reached higher accuracies than the frequency-domain set because of fewer intercorrelations. An increase in the number of neighbours for k-NN used on complete sets lessens the model accuracy because class boundaries overlap and are relatively noisy. Above the k-value of thirteen, the trend for complete feature sets of accuracy decrease flattens out gradually. Predictors aggregated from tri-axis recording performed better in classification than those from one axis.

The k-NN accuracy with different feature subset combinations pulled out of the complete sets has a standard deviation of 7\%, and it is 10\% within IQR. The feature selection methods can pick, in most cases, a group with accuracy above the upper quartile of accuracy distribution of every attribute combination with the same cardinality. Training k-NN on three principal components produced worse accuracy than filtering feature selection.

The ensemble feature selection methods as the rank product are necessary to balance the adverse effect of individual selection metrics. In the slight majority of 44\% of 216 dataset alterations, the rank product found the best-performing attributes closely followed by mutual information. The rank product also reaches the 97.50\% percentile in the accuracy distribution of the same-sized sets. Mutual information has a better distribution percentile by 2.84\% than the rank product \textbf{(RQ3)}.

The number of features has a direct proportionality effect accuracy but at most 6\% increase is observed while moving from two to three features. The absolute best predictor triplets for MaFaulDa are in TD (accuracy of 85.47\%): zero-crossing rate, peak-to-peak, skewness, and in FD (accuracy of 87.52\%): spectral centroid, roll-off frequency, and entropy \textbf{(RQ3)}.

@@

Collected dataset analysis confirmed the stationarity of the signal spectrum and the five-second long time series as satisfactory for feature extraction. The dataset serves as a practical demonstration of labeling potentially future faults from bearing defect frequency amplitudes. It also proved the immense difficulty of obtaining observations of faulty machinery parts in the field without purposefully damaging bearings and offsetting the shaft \textbf{(RQ4)}.

@@

\textbf{Future work} can extend various aspects and answer additional questions arising from the experiments in the thesis. The natural extension is to confirm that the results hold in other publicly available datasets of machinery faults, e.g. in CWRU. The other alteration is to use more and different starting base feature sets. Harmonic frequencies or wavelet coefficients can generate predictors that could be further reduced to the smallest possible subsets.

Ensemble strategies inspired by classical election systems may be utilized in the task of feature selection. Different types of classifiers for online learning could be compared to k-nearest neighbours, e.g. Hoeffding tree. A data logger could be combined with the classifier model and a mechanism to annotate observation either automatically depending on the overall vibration level or manually during maintenance or after an unexpected failure.

The strategic aim would be to connect sensors on a large scale on multiple test rigs to train individual models on each bearing and combine them in, for example, a federative learning approach. Of course, vibration monitoring can be employed in other diagnostic tasks without direct expert knowledge, such as train wheel or rail degradation due to wear and tear.
